---
title: "Mellon Progress Report Elements"
author: "Jamison Crawford, MPA"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    theme: readable
    df_print: paged
    highlight: tango
    toc: yes
    toc_float: yes
---

```{r include = FALSE, cache = TRUE}

knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE)

setwd("~/CASA HIP")

load(file = "hip_final.RDATA")

options(scipen = 999)

library(kableExtra)

```

<br>
<br>

# Overview

The following provides data tables, visualizations, descriptions, salient preprocessing
steps, and data nuances for the **Mellon Interim Progress Report**. Elements are provided 
separately for streamlined report inclusion.

<br>
<br>

## Reproducible Documentation

All documentation to reproduce preprocessing and the present reporting elements, in their 
entirety, are available in 
[**this GitHub repository**](https://github.com/jamisoncrawford/hip/blob/master/README.md).

<br>
<br>

## Information Security

All data herein are summarized and exclude personally-identifiable information. However, 
because **Humanities Inclusivity Program** (**HIP**) scholars are few, particularly when 
parsed by calendar year (some of which have one to two HIP scholars), these observations 
**should be treated as identifiable**.

**Disclaimer:** The Institutional Researcher and author of the present report recommends 
suppression of all data barring program-wide summary statistics to protect HIP scholars' 
identities and performance in accordance with the **Family Educational Rights & Privacy** 
**Act** (**FERPA**).

<br>
<br>

## Data Nuances

Of the sixteen **HIP** scholars, one scholar has been removed from analysis for 
matriculation year and performance-related outliers, resulting in fifteen total scholars.

For comparison, only **non-HIP** students active in the same calendar years and enrolled 
in the same majors are included (n = 14,123). Only courses common to **HIP** and 
**non-HIP** students are used (n = 1,294,170).

<br>
<br>

# Graphics

The following data visualizations are accompanied by descriptions and data nuances.

<br>
<br>

## Class-Course Level Differences

`Student Class Level v. Course Level` depicts the total proportion of courses 
for **HIP** and **non-HIP** students that are at, above, or below student class 
(i.e. `Freshman`, `Sophomore`, `Junior`, or `Senior`). 

Course levels are determined by the first digit of each course number, e.g. `1010` 
indicates a `1-` or `1000-` level course.

Class levels are determined by the BoR-designated `Student Level` in the GSU Data 
Warehouse. For example, `10` indicates `Freshman`, `20` indicates `Sophomore`, etc.

<br>

```{r fig.align = "center", fig.width = 8, fig.height = 4, cache = TRUE}

viz1

```

<br>
<br>

## Mean GPA v. Mean Course Levels

`Student Class Level v. Course Level` depicts the mean `GPA` of all **HIP** and 
**non-HIP** against mean class-course levels, i.e. the difference between 
BoR-designated student levels and each course number category.

Where n > 2, linear models are provided depicting mean `GPA` as a function of 
mean class-course difference.

<br>

```{r fig.align = "center", fig.width = 8, fig.height = 5, cache = TRUE}

viz2

```

<br>
<br>

## Mean GPA by Calendar Year

`Mean GPA by Calendar Year` depicts the mean `GPA` of all **HIP** and 
**non-HIP** over calendar years. All mean `GPA` values are "jittered" 
vertically and horizontally (i.e. white noise is added to coordinate mapping). 

Notably, **HIP** scholars' mean `GPA` values are only jittered horizontally. In 
other words, they accurately reflect the mean `GPA` values indicated by the 
y-axis. 

<br>

```{r fig.align = "center", fig.width = 8, fig.height = 4, cache = TRUE}

viz3

```

<br>
<br>

## Mean GPA by Year & Class

`Mean GPA by Calendar Year & Class` depicts a time series of the mean `GPA` of 
all **HIP** and **non-HIP** students over calendar years and faceted by 
BoR-designated class (i.e. `Freshman`, `Sophomore`, etc.).

**Emphasis:** While **HIP** scholars tend to have higher mean `GPA` values than 
**non-HIP** students, there is an exception among `Senior` class scholars. 
However, the 2019 mean `GPA` values buck this trend, potentially indicating 
improved **HIP** programming.

<br>

```{r fig.align = "center", fig.width = 8, fig.height = 4, cache = TRUE}

viz4

```

<br>
<br>

# Tables

The following provides tabular data on the most highly-aggregated groups 
comparing **HIP** scholars and **non-HIP** students.

<br>
<br>

## Course-Class Differences

`Course-Class Differences` provides **HIP** and **non-HIP** students as described 
in `Student Class Level v. Course Level`. Percent differences are included for 
enhanced granularity.

<br>

```{r, cache = TRUE}

lvls_diff_tbl %>%
  kable() %>%
  kable_styling(bootstrap_options = c("hover", "striped"))

```

<br>
<br>

## Annual Course-Class Difference

`Annual Course-Class Difference` provides **HIP** and **non-HIP** students as 
described in `Student Class Level v. Course Level`, albeit parsed by year. 
Percent differences are included for enhanced granularity.

<br>

```{r, cache = TRUE}

lvls_diff_yr_tbl %>%
  kable() %>%
  kable_styling(bootstrap_options = c("hover", "striped"))

```

<br>
<br>

## Summary GPA by Class

`Summary GPA by Class` provides **HIP** and **non-HIP** student mean and median 
`GPA` values aggregated by class (e.g. `Freshman`, `Sophomore`, etc.). Percent 
differences are included for enhanced granularity.

<br>

```{r, cache = TRUE}

gpa_mean_median_class_tbl %>%
  kable() %>%
  kable_styling(bootstrap_options = c("hover", "striped"))

```

<br>
<br>

## Summary GPA by Class & Year

`Summary GPA by Class & Year` provides **HIP** and **non-HIP** student mean and 
median `GPA` values aggregated by class (e.g. `Freshman`, `Sophomore`, etc.) and 
parsed by calendar year. Percent differences are included for enhanced granularity.

<br>

```{r, cache = TRUE}

gpa_mean_median_class_yr_tbl %>%
  kable() %>%
  kable_styling(bootstrap_options = c("hover", "striped"))

```

<br>
<br>

## Summary GPA by Course & Year

`Summary GPA by Course & Year` provides **HIP** and **non-HIP** student mean and 
median `GPA` values aggregated by class (e.g. `Freshman`, `Sophomore`, etc.),  
parsed by calendar year and course level. Percent differences are included for 
enhanced granularity.

<br>

```{r, cache = TRUE}

regc_mean_med_tbl %>%
  kable() %>%
  kable_styling(bootstrap_options = c("hover", "striped"))

```

<br>
<br>